---
title: "Unit9HW"
author: "Quentin Thomas"
date: "3/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MSDS 6306 Preparing Data

### Questions

1. **Harry Potter Casts**


`a.` In the IMDB, there are listings of full cast members for movies. Navigate to [link](http://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1). Feel free to View Source to get a good idea of what the page looks
like in the code.

`b.` Scrape the page with any R package that makes things easy for you. Of particular interest is the table of the Cast in order of
crediting. Please scrape this table (You might have to fish it out of several of the tables from the page) and make it a data.frame() of the Cast in your R environment.

```{r}
library(rvest)

page_data <- read_html(x = "http://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1")
page_table <- html_node(page_data, xpath = "/html/body/div[1]/div/div[4]/div[3]/div[1]/div[1]/div[2]/table[3]")
casts <- html_table(page_table)
casts
```

`c.` clean up the table

- It should not have blank observations or rows, a row that should be column names, or just '...'

```{r}
library(stringr)
casts <- casts[, c(2,4)]
casts <- casts[-1,]
casts$X4 <- str_replace_all(casts$X4, "[\r\n]" , "")
```

- It should have intuitive column names (ideally 2 to start - Actor and Character)
```{r}
names(casts) <- c("Actor", "Character")
```

- In the film, Mr Warwick plays two characters which makes his row look a little weird. Please replace his character column with just
"Griphook/Professor Filius Flitwick" to make it look better

```{r}
casts[grepl(pattern = "^Warwick", casts$Actor),]$Character <- "Griphook/Professor Filius Flitwick"
```

- One row might result in "Rest of cast listed alphabetically" - remove this observation.

```{r}
casts <- casts[!c(casts$Actor == "Rest of cast listed alphabetically:"),]
casts
```

`d.` Split the Actor's name into two columns: FirstName and Surname. Keep in mind that some actors/actresses have middle names as well. Please make sure that the middle names are in the FirstName column, in addition to the first name(example: given the Actor Frank Jeffrey Stevenson, the Firstname column would say "Frank Jeffrey.")


```{r}
#Need to split off the last names from the rest of the names so that it will also include middle names if they have it.
split_names <- str_split(casts$Actor, pattern = " (?=[^ ]+$)")
d <- data.frame(matrix(unlist(split_names), nrow=149, byrow=T),stringsAsFactors=FALSE)
names(d) <- c("FirstName","Surname")
casts <- cbind(d, casts$Character)
names(casts)[1] <- "FirstName"
names(casts)[3] <- "Character"
```

`e.` Present the first 10 rows of the data.frame()- It should have only FirstName, Surname, and Character Columns.

```{r}
head(casts, 10)
```

2. **SportsBall**

`a.` On the ESPN website, there are statistics of each NBA player. Navigate to the San Antonio Spurs current statistics [link](http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs). You are interested in the `Shooting Statistics` table.

`b.` Scrape the page with any R package that makes things easy for you. There are a few tables on the page, so make sure you are targeting specifically the Shooting
statistics table.
```{r}
#scraping all the data!
espn_data <- read_html(x = "http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs")
espn_table <- html_node(espn_data, css ="#my-players-table > div.mod-container.mod-table > div:nth-child(3) > table")
shooting_stats <- html_table(espn_table)
shooting_stats
```

`c.` Clean up the table (You might get some warnings if you're working with tibbles)
- You'll want to create an R data.frame() with one observation for each player. Make sure that you do not accidentally include blank rows, a row of column
names, or the Totals row in the table as observations.

```{r}
#Remove SHooting Stats CSS nonsense that is cluttering up my dataframe
shooting_stats <- shooting_stats[-1,]
#Replacing the label generic x labels with the actual header values as the column names
colnames(shooting_stats) <-as.character(unlist(shooting_stats[1,]))
shooting_stats <- shooting_stats[-1,]
#Removing totals row
shooting_stats <- shooting_stats[-nrow(shooting_stats),]
shooting_stats
```


- The Column PLAYER has two variables of Interests in it: the player's name and their position, denoted by 1-2 letters after their name. Split the cells into
two columns, one with Name and the other Position.

```{r}
names_and_pos <- str_split(shooting_stats$PLAYER, ", ")
df <-  data.frame(matrix(unlist(names_and_pos), nrow=17, byrow=T),stringsAsFactors=FALSE)
names(df) <- c("Player", "Position")
df
stats <- shooting_stats[, c(2:15)]
sa_playerstats <- cbind(df, stats)
sa_playerstats
```

- Check the data type of all columns. convert relevant columns to numeric. Check the data type of all columns again to confirm that they have changed!

Data types of relevant columns BEFORE conversion.
```{r}
sapply(sa_playerstats, mode)
```


Data types of relevant columns After conversion

```{r}
stats <- as.data.frame(sapply(sa_playerstats[,c(3:16)], as.numeric))
sa_playerstats <- cbind(df, stats)
sapply(sa_playerstats, mode)
```

`d.` Create a colorful bar chart that shows the Field Goals Percentage per game for each person. It will be graded on the following criteria.

- Informative Title, centered
- Relevant x and y axis labels (not simply variables names!)
- Human-readable axes with no overlap(you might have to flip x and y to fix that)
- Color the columns by the team member's position (so, all PF's should have the same color, etc.)

```{r}
library(ggplot2)

sports_theme <- theme(plot.title = element_text(face = "bold.italic",
                                                size= 14,
                                                color = "black"),
                      axis.text.x = element_text(size = 8, face = "bold.italic"),
                      axis.text.y = element_text(face = "bold.italic"),
                      legend.text = element_text(face = "bold.italic", color = "black", size = 8),
                      legend.title = element_text(face = "bold", color ="black", size = 14),
                      axis.title.x = element_text(face= "bold", color = "black", size = 14),
                      axis.title.y = element_text(face="bold", color = "black", size = 14))

ggplot(data = sa_playerstats, aes(x = Player, y = `FG%`, fill = Position)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
  coord_flip() +
  labs(title = "San Antonio Spurs Field Goal Percentage 2017-2018") +
  sports_theme


```