---
title: "MSDS 6306 HW 4"
author: "Quentin Thomas"
date: "January 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Questions

### 1 Navigate on GitHub to https://github.com/rudeboybert/fivethirtyeight and read README.md. Seriously, it will include everything you need. 

  `a)` Install the `fivethrityeight` package

```{r}

#load the library 
library(fivethirtyeight)
```

  `b)` In the listing of data sets in package `fivethirtyeight` assign the 18th data set to a object called `df`
  
```{r}
# view Datasets
data(package = "fivethirtyeight")

# 18th data set is college_recent_grads

df <- college_recent_grads
```

  `c)` Use a more detailed list of the data sets to write out the URL in a comment to the related news story.

```{r}
# https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/
```

  `d)` Using R command(s), give the dimensions and column names of this data frame
  
```{r}
dim(df)
```

### 2 Use your newly assigned dataframe for Question 2

  `a)` Write an R command that gives you the column names of the data frame.  Right after that, write one that counts the number of columns but not rows.  Hint: The number should match one of your       numbers in Question 1d for dimensions.
  
```{r}
colnames(df)

ncol(df)

```

  `b` Generate a count of each unique major_category in the data frame.  I recommend using libraries to help.  I have demonstrated one briefly in live-session.  To be clear, this should look like a matrix or data frame containing the major_category and the frequency it occurs in the dataset.  Assign it to major_count.
  
```{r}
library(vcd)
major_counts <- table(df$major_category)
major_counts

```

  `c`	To make things easier to read, put par(las=2) before your plot to make the text perpendicular to the axis.  Make a barplot of major_count.  Make sure to label the title with something informative (check the vignette if you need), label the x and y axis, and make it any color other than grey.  Assign the major_category labels to their respective bar.  Flip the barplot horizontally so that bars extend to the right, not upward.  All of these options can be done in a single pass of barplot(). Note: It’s okay if it’s wider than the preview pane.
  
```{r}
par(mar=c(10,12,4,2))
par(las=2)
barplot(major_counts,
        main = "College Recent Grads by Major",
        horiz = TRUE,
        xlab = "Number of Graduates",
        sub = "Fivethrityeight data", col="purple",
        cex.lab = 1.5,
        names.arg = names(major_counts),
        cex.names=0.8)
```

  `d` Write the fivethirtyeight data to a csv file.  Make sure that it does not have row labels.
  
```{r}
write.csv(df, file = "college_recent_grads.csv", row.names=FALSE, na="")
```

### 3 Cookbook

  `a` Start a new repository on GitHub for your SMU MSDS homework.  On your local device, make sure there is a directory for Homework at the minimum; you are welcome to add whatever you would like to this repo in addition to your requirements here.
  
```bash
git init msds_6306
mkdir hw
cd hw
```

  `b` Create a README.md file which explains the purpose of the repository, the topics included, the sources for the material you post, and contact information in case of questions. Remember, the one in the root directory should be general.  You are welcome to make short READMEs for each assignment individually in other folders.
  
```bash
touch README.md
```

  `c` In one (or more) of the nested directories, post your RMarkdown script, HTML file, and data from ‘fivethirtyeight.’  Make sure that in your README or elsewhere that you credit fivethirtyeight in some way.
  
```bash
#adds all the files that are in directory
git add .
```

  `d` In your RMarkdown script, please provide the link to this GitHub so the grader can see it.
  
  [Link To Github for HW 4](https://github.com/TheQuengineer/msds_6306/hw/unit_4)